Clustering in a hyperbolic model of complex networks.

### 2019-04-02 proofs_finite_k.tex
implementation of Nikolaos comments in section 4.

Due to the coupling, the KPKVB graph or its poissonization is contained in the other, i.e. we can use new graph labels $G_1 \subset G_2$; e.g. if $N\geq n$, then $G_1 = G(n;\alpha,\nu)$ and $G_2 = \GPo(n;\alpha,\nu)$. We write $V_{1,k}$ for the set of vertices of $G_1$ which have $k$ neighbours in $G_1$ and $V_{2,k}$ for the set of vertices of $G_2$ which have $k$ neighbours in $G_2$. Write $N_{1,k}=|V_{1,k}|$ and $N_{2,k}=|V_{2,k}|$.

@@

With this notation, we can rewrite:
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &= \left|\sum_{v \in V_{1,k}}\frac{c_{G_1}(v)}{N_{1,k}}-\sum_{v \in V_{2,k}}\frac{c_{G_2}(v)}{N_{2,k}}\right| \\
&=\left|\sum_{v \in W_1}\frac{c_{G_1}(v)}{N_{1,k}}+\sum_{v \in W_2}\frac{c_{G_1}(v)}{N_{1,k}}-\sum_{v \in W_1}\frac{c_{G_2}(v)}{N_{2,k}}-\sum_{v \in W_3}\frac{c_{G_2}(v)}{N_{2,k}}\right|
\end{align*}

For every vertex of $W_1$, the clustering coefficient in $G_1$ and in $G_2$ have the same value. Hence, we can simplify:
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &= \left|\sum_{v \in W_1}c_{G_1}(v)\frac{N_{2,k}-N_{1,k}}{N_{1,k}N_{2,k}}+\sum_{v \in W_2}\frac{c_{G_1}(v)}{N_{1,k}}-\sum_{v \in W_3}\frac{c_{G_2}(v)}{N_{2,k}}\right|
\end{align*}
Using the triangle inequality for the absolute value and the fact that $|c_G(v)|\leq 1$,
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &\leq \frac{|W_1| |N_{2,k}-N_{1,k}|}{N_{1,k}N_{2,k}}+\frac{|W_2|}{N_{1,k}}+\frac{|W_3|}{N_{2,k}}
\end{align*}
Using that $|W_1|\leq N_{1,k}$,
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &\leq \frac{|N_{2,k}-N_{1,k}|}{N_{2,k}}+\frac{|W_2|}{N_{1,k}}+\frac{|W_3|}{N_{2,k}}
\end{align*}

@@

By the paper by Gugelmann et al.~\cite{gugelmann2012random} it holds that $N_{1,k}$ or $N_{2,k}$ is $\Theta(n)$ a.a.s. for fixed degree $k$ and as $|N_{1,k}-N_{2,k}| =O(\sqrt{n\log n})$ a.a.s., it follows that both $N_{1,k},N_{2,k} = \Theta(n)$ a.a.s. Putting everything together yields the claim:
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &\leq \frac{|N_{2,k}-N_{1,k}|}{N_{2,k}}+\frac{|W_2|}{N_{1,k}}+\frac{|W_3|}{N_{2,k}} \xrightarrow[n\rightarrow\infty]{\Pee} 0
\end{align*}

@@

\subsubsection{Truncation}
Consider the truncated Poissonized KPKVB model $\GPoH$ with $H =H_n= \frac{R}{4}$.
%H,R-H tend to infinity and coupling is still exact
\begin{lemma}\label{lem:n1}
Let $N_1(H)$ denote the number of vertices with radial coordinate at most $R-H \in [0,R]$ in the Poissonized KPKVB model $\GPo$. If $H \rightarrow \infty$, then $\E N_1(H) = o(n)$.
%For all $\epsilon >0$, there is $H_\epsilon$ such that for all $H\geq H_\epsilon$, it holds that $\E N_1(H) \leq \epsilon n$.
\end{lemma}
\begin{proof}
By the Campbell-Mecke formula applied to the vertex set of $\GPo$, by the asymptotic behaviour of $\cosh$ and using that all $R, R-H, H \rightarrow \infty$,
\begin{align*}
\E N_1(H) = n \frac{\cosh(\alpha(R-H))-1}{\cosh(\alpha R) -1} = (1+o(1)) n e^{-\alpha H} = o(n)
\end{align*}
\end{proof}
\begin{lemma}\label{lem:n2}
Let $N_2(H)$ denote the number of vertices with at least one neighbour with radial coordinate at most $R-H \in [0,R]$ in the Poissonized KPKVB model. If $H \rightarrow \infty$, then $\E N_2(H) = o(n)$.
\end{lemma}
\begin{proof}
Among the vertices with at least one neighbour with radial coordinate at most $R-H$, we distinguish between those that themselves have radial coordinate at most $R-H_0$ and those that themselves have radial coordinate at least $R-H_0$ where $H_0 = \log H$. We will show that the expectation of both numbers is sublinear in $n$.

If $H \rightarrow \infty$, then also $H_0 = \log H \rightarrow \infty$. Hence by Lemma~\ref{lem:n1}, the expected number of vertices with radial coordinate at most $R-H_0$ (in particular those vertices among them with at least one neighbour with radial coordinate at most $R-H$), $\E N_1(H_0) = o(n)$ is sublinear. For a vertex with radial coordinate at least $R-H_0$, the probability that it has a neighbour with radial coordinate at most $R-H$, tends to zero. Therefore, by the Campbell-Mecke formula the number of such vertices is sublinear.
\end{proof}

@@

Using a telescoping sum, we obtain that
\begin{align*}
|c(k;\GPo)-c(k;\GPoH)| \leq & \left|c(k;\GPo)-\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_n(v)\right|\\
&+\left|\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_n(v) -\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)}c_{n,H}(v)\right| \\
&+\left|\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_{n, H}(v)-c(k;\GPoH)\right| 
\end{align*}
Now we use $X_1 := |V_n(k)\backslash V_{n,H}(k)|$, $X_2 := \sum_{v \in V_{n,H}(k)}|c_n(k)-c_{n,H}(k)|$ and $X_3 :=|N_{n,H}(k)-N_n(k)|$ to bound the three differences. In the first difference, after factorizing $N_n(k)$, all terms in $V_{n,H}(k)$ cancel and the remaining terms are all bounded by one:
\begin{align*}
\left|c(k;\GPo)-\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_n(v)\right| = \left| \frac{1}{N_n(k)}\sum_{v \in V_n \backslash V_{n,H}(k)} c_n(v) \right| \leq \frac{X_1}{N_n(k)}
\end{align*}
In the second difference, the sums can be aligned and then the triangle inequality for the absolute value applied:
\begin{align*}
\left|\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_n(v) -\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)}c_{n,H}(v)\right| \leq \frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} \left|c_n(v)-c_{n,H}(v)\right| =\frac{X_2}{N_n(k)}
\end{align*}
In the third difference, we can simplify:
\begin{align*}
\left|\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_{n, H}(v)-c(k;\GPoH)\right| &= \left|\sum_{v\in V_{n,H}(k)} c_{n,H} \left(\frac{1}{N_n(k)}-\frac{1}{N_{n,H}(k)}\right)\right| \\
&= \frac{|N_{n,H}(k)-N_n(k)|}{N_n(k)N_{n,H}(k)}\sum_{v \in V_{n,H}(k)}c_{n,H}(k)
\end{align*}
Thus, we end up with the upper bound, using that $c_{n,H}(k) \leq 1$ and $N_{n,H}(k) = |V_{n,H}(k)|$,
\begin{align*}
|c(k;\GPo)-c(k;\GPoH)| \leq \frac{X_1}{N_n(k)}+\frac{X_2}{N_n(k)}+\frac{X_3}{N_n(k)}\frac{\sum_{v \in V_{n,H}(k)}c_{n,\leq H}(k)}{N_{n,H}(k)} \leq \frac{X_1+X_2+X_3}{N_n(k)}
\end{align*}
Now using that a.a.s. $N_n(k) \geq a_1(k)n$ gives that a.a.s.
\begin{align*}
|c(k;\GPo)-c(k;\GPoH)| \leq \frac{X_1+X_2+X_3}{a_1(k)n} 
\end{align*}

By Lemma~\ref{lem:n1}, we have that $\E X_1 = \E X_3 = o(n)$ and by Lemma~\ref{lem:n2}, we have that $\E X_2 = o(n)$. We conclude that $\E |c(k;\GPo)-c(k;\GPoH)| = o(1)$. From this, convergence in probability follows by Markov's inequality: for $\epsilon >0$, $\Pee(|c(k;\GPo)-c(k;\GPoH)| \geq \epsilon) \leq \frac{\E|c(k;\GPo)-c(k;\GPoH)|}{\epsilon} = o(1)$.
\end{proof}
\subsubsection{Coupling}
\begin{lemma}\label{lem:coup}
The clustering function in the truncated Poissonized KPKVB model $\GPoH$ and in the truncated box model $\GboxH$ agree, i.e. $c(k;\GPoH) = c(k;\GboxH)$.
\end{lemma}
\begin{proof}
As all vertices involved have height $\leq H$ resp. radial coordinate $\geq R-H$ and as $H\leq \frac{R}{4}$, all vertices involved (in the disk)  will have radial coordinate $\geq \frac{3}{4}R$, such that by Lemma~30(ii) from the paper on the largest component of KPKVB, see Lemma~\ref{lem:coupling_edges}, edges exist between such vertices in the disk if and only if they exist in the box. So the two graphs are isomorphic and hence the clustering functions agree.
\end{proof}

@@

By the Campbell-Mecke formula, we have
\begin{align*}
\E X = \frac{1}{n}\int_{\Rcal_H} \E h(V,v) \mu(dv) = \int_0^H \E h(V,(0,y))\alpha e^{-\alpha y} dy
\end{align*}
Then, we condition on that the additional vertex $(0,y) \in \Rcal_H$ has degree $k$,
\begin{align*}
\E X = \int_0^H \E[h(V,(0,y))|D_{G_{box,H}}(0,y)=k]\Pee(D_{G_{box,H}}(0,y)=k)\alpha e^{-\alpha y}dy
\end{align*}
We observe that for $H \rightarrow \infty$, the integrand converges point-wise for $y \in (0,\infty)$ to $P(y)\rho(y,k)\alpha e^{-\alpha y}$ and is always bounded by the integrable function $\alpha e^{-\alpha y}$.
Therefore, by the theorem of dominated convergence, as $H \rightarrow \infty$,
\begin{align*}
\E X \rightarrow \int_0^\infty P(y)\rho(y,k)\alpha e^{-\alpha y}dy
\end{align*}
Note that for $X^2$, the summation can be split into off-diagonal (with $v_1 \not = v_2$) and diagonal (with $v_1=v_2$) terms, i.e. write $X^2 = \frac{1}{n^2}(Y_1+Y_2)$ where 
$$Y_1 = \sum_{v_1 \not = v_2 \in V(G_{box,H})} h(V(G_{box,H}),v_1)h(V(G_{box,H}),v_2)$$ and $Y_2 = \sum_{v_1 \in V(G_{box,H})} h(V(G_{box,H}),v_1)^2$. As $h \leq 1$, $Y_2 \leq n X\leq n$ and so, $\E Y_2 \leq n$, we have by the Campbell-Mecke formula,
\begin{align*}
\E Y_1 = \int_{\Rcal_H} \int_{\Rcal_H} \E[h(V(G_{box,H}),v_1)h(V(G_{box,H}),v_2)]\mu(dv_1)\mu(dv_2)
\end{align*}
If $v_1=(x_1,y_1)$ and $v_2=(x_2,y_2)$ have horizontal distance $>2e^H$, their neighbourhood balls are disjoint w.r.t. the connection rule of the truncated box model $\GboxH$. To see this, assume that $v_3=(x_3,y_3)$ is adjacent to both $v_1$ and $v_2$, then $|x_1-x_3|\leq e^{\frac{1}{2}(y_1+y_3)}$ and $|x_2-x_3|\leq e^{\frac{1}{2}(y_2+y_3)}$. As $y_1, y_2, y_3 \leq H$, it follows that $|x_1-x_3| \leq e^H$ and $|x_2-x_3|\leq e^H$, which implies that $|x_1-x_2| \leq |x_1-x_3|+|x_2-x_3|\leq 2e^H$, contradiction.

Due to the disjointedness of the neighbourhood balls and the properties of the Poisson process $h(V,v_1)$ and $h(V,v_2)$ are independent. The horizontal distance is uniform in $[0,\frac{\pi n}{2\nu}]$. Therefore, the probability that two vertices have horizontal distance $\leq 2e^H$ is $\leq \frac{4\nu e^H}{\pi n} = o(1)$. Writing $d_H$ for the horizontal distance, we can continue the calculation of the second moment,
\begin{align*}
\E Y_1 &= \int_{} \int_{\substack{\Rcal_H\times \Rcal_H:\\ |x_1-x_2|>2 e^H}} \E[ h(V,v_1)]\E[h(V,v_2)]\mu(dv_1)\mu(dv_2)\\
&\qquad+\int_{} \int_{\substack{\Rcal_H\times \Rcal_H:\\ |x_1-x_2|\leq2 e^H}} \E[ h(V,v_1)]\E[h(V,v_2)]\mu(dv_1)\mu(dv_2) \\
&= (1+o(1))\int_{} \int_{\Rcal_H\times \Rcal_H} \E[ h(V,v_1)]\E[h(V,v_2)]\mu(dv_1)\mu(dv_2) \\
&= (1+o(1)) (n\E X)^2%+o((n\E X)^2)+o(n)
\end{align*}
It follows that $\E X^2 = (1+o(1))(\E X)^2$, which implies $Var(X) \rightarrow 0$. We conclude
$$X \xrightarrow[n\rightarrow\infty]{\Pee} \int_0^\infty P(y)\rho(y,k)\alpha e^{-\alpha y}dy.$$
In a similar way, by the Campbell-Mecke formula,
\begin{align*}
\E Z =\frac{1}{n} \int_{\Rcal_H} \Pee(D_{G_{box,H}}(v)=k) \mu(dv) = \int_0^H \Pee(D_{G_{box,H}}(0,y)=k) \alpha e^{-\alpha y}dy
\end{align*}
By the theorem of dominated convergence, it follows that
\begin{align*}
\E Z \rightarrow \int_0^\infty \rho(y,k)\alpha e^{-\alpha y}dy
\end{align*}
The statement that $Var(Z) \rightarrow 0$ follows by the same steps as for $X$, just using $h(V,v) = \1_{\{D_{\GboxH}(v)=k\}}$ instead. Hence,
\begin{align*}
\E Z \xrightarrow[n\rightarrow\infty]{\Pee} \int_0^\infty \rho(y,k)\alpha e^{-\alpha y}dy
\end{align*}
Finally note that 
\begin{align*}
\gamma(k) = \frac{\int_0^\infty P(y)\rho(y,k)\alpha e^{-\alpha y}dy}{\int_0^\infty \rho(y,k)\alpha e^{-\alpha y}dy}.
\end{align*}

>>>>

Due to the coupling, the KPKVB graph or its poissonization is contained in the other, i.e. we can use new graph labels $G_1 \subset G_2$; e.g. if $N\geq n$, then $G_1 \stackrel{d}{=} G(n;\alpha,\nu)$ and $G_2 \stackrel{d}{=} \GPo(n;\alpha,\nu)$. We write $V_{1,k}$ for the set of vertices of $G_1$ which have $k$ neighbours in $G_1$ and $V_{2,k}$ for the set of vertices of $G_2$ which have $k$ neighbours in $G_2$. Write $N_{1,k}=|V_{1,k}|$ and $N_{2,k}=|V_{2,k}|$.

@@

With this notation, we can rewrite:
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &= \left|\sum_{v \in V_{1,k}}\frac{c_{G_1}(v)}{N_{1,k}}-\sum_{v \in V_{2,k}}\frac{c_{G_2}(v)}{N_{2,k}}\right| \\
&=\left|\sum_{v \in W_1}\frac{c_{G_1}(v)}{N_{1,k}}+\sum_{v \in W_2}\frac{c_{G_1}(v)}{N_{1,k}}-\sum_{v \in W_1}\frac{c_{G_2}(v)}{N_{2,k}}-\sum_{v \in W_3}\frac{c_{G_2}(v)}{N_{2,k}}\right|.
\end{align*}

For every vertex of $W_1$, the clustering coefficient in $G_1$ and in $G_2$ have the same value. Hence, we can simplify:
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &= \left|\sum_{v \in W_1}c_{G_1}(v)\frac{N_{2,k}-N_{1,k}}{N_{1,k}N_{2,k}}+\sum_{v \in W_2}\frac{c_{G_1}(v)}{N_{1,k}}-\sum_{v \in W_3}\frac{c_{G_2}(v)}{N_{2,k}}\right|.
\end{align*}
Using the triangle inequality for the absolute value and the fact that $|c_G(v)|\leq 1$,
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &\leq \frac{|W_1| |N_{2,k}-N_{1,k}|}{N_{1,k}N_{2,k}}+\frac{|W_2|}{N_{1,k}}+\frac{|W_3|}{N_{2,k}}.
\end{align*}
Using that $|W_1|\leq N_{1,k}$,
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &\leq \frac{|N_{2,k}-N_{1,k}|}{N_{2,k}}+\frac{|W_2|}{N_{1,k}}+\frac{|W_3|}{N_{2,k}}.
\end{align*}

@@

By the paper by Gugelmann et al.~\cite{gugelmann2012random} it holds that $N_{1,k}$ or $N_{2,k}$ is $\Theta(n)$ a.a.s. for fixed degree $k$ and as $|N_{1,k}-N_{2,k}| =O(\sqrt{n\log n})$ a.a.s., it follows that both $N_{1,k},N_{2,k} = \Theta(n)$ a.a.s. Putting everything together yields the claim:
\begin{align*}
|c(k;G_n)-c(k;\GPo)| &\leq \frac{|N_{2,k}-N_{1,k}|}{N_{2,k}}+\frac{|W_2|}{N_{1,k}}+\frac{|W_3|}{N_{2,k}} \xrightarrow[n\rightarrow\infty]{\Pee} 0.
\end{align*}

@@

\subsubsection{Truncation}
Consider the truncated Poissonized KPKVB model $\GPoH$ with $H =H_n= \frac{R}{4}$.
%H,R-H tend to infinity and coupling is still exact
\begin{lemma}\label{lem:n1}
Let $N_1(H)$ denote the number of vertices with radial coordinate at most $R-H \in [0,R]$ in the Poissonized KPKVB model $\GPo$. Then $\E N_1(H) = o(n)$.%If $H \rightarrow \infty$, 
%For all $\epsilon >0$, there is $H_\epsilon$ such that for all $H\geq H_\epsilon$, it holds that $\E N_1(H) \leq \epsilon n$.
\end{lemma}
\begin{proof}
By the Poisson distribution of the number of points inside the disk with radius $R-H$, by the asymptotic behaviour of $\cosh$ and using that all $R, R-H, H \rightarrow \infty$,
\begin{align*}
\E N_1(H) = n \frac{\cosh(\alpha(R-H))-1}{\cosh(\alpha R) -1} = (1+o(1)) n e^{-\alpha H} = o(n)
\end{align*}
\end{proof}
\begin{lemma}\label{lem:n2}
Let $N_2(H)$ denote the number of vertices with at least one neighbour with radial coordinate at most $R-H \in [0,R]$ in the Poissonized KPKVB model. If $H \rightarrow \infty$, then $\E N_2(H) = o(n)$.
\end{lemma}
\begin{proof}
Among the vertices with at least one neighbour with radial coordinate at most $R-H$, we distinguish between those that themselves have radial coordinate at most $R-H_0$ and those that themselves have radial coordinate at least $R-H_0$ where $H_0 = \log H$. We will show that the expectation of both numbers is sublinear in $n$.

Since $H \rightarrow \infty$ we also have that $H_0 = \log H \rightarrow \infty$. Hence by Lemma~\ref{lem:n1}, the expected number of vertices with radial coordinate at most $R-H_0$ (in particular those vertices among them with at least one neighbour with radial coordinate at most $R-H$), $\E N_1(H_0) = o(n)$ is sublinear. For a vertex with radial coordinate at least $R-H_0$, the probability that it has a neighbour with radial coordinate at most $R-H$, tends to zero. Therefore, by the Campbell-Mecke formula the number of such vertices is sublinear.
\end{proof}

@@

Using a telescoping sum, we obtain that
\begin{align*}
|c(k;\GPo)-c(k;\GPoH)| \leq & \left|c(k;\GPo)-\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_n(v)\right|\\
&+\left|\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_n(v) -\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)}c_{n,H}(v)\right| \\
&+\left|\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_{n, H}(v)-c(k;\GPoH)\right|. 
\end{align*}
Now we use $X_1 := |V_n(k)\backslash V_{n,H}(k)|$, $X_2 := \sum_{v \in V_{n,H}(k)}|c_n(k)-c_{n,H}(k)|$ and $X_3 :=|N_{n,H}(k)-N_n(k)|$ to bound the three differences. In the first difference, after factorizing $N_n(k)$, all terms in $V_{n,H}(k)$ cancel and the remaining terms are all bounded by one:
\begin{align*}
\left|c(k;\GPo)-\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_n(v)\right| = \left| \frac{1}{N_n(k)}\sum_{v \in V_n \backslash V_{n,H}(k)} c_n(v) \right| \leq \frac{X_1}{N_n(k)}.
\end{align*}
In the second difference, the sums can be aligned and then the triangle inequality for the absolute value applied:
\begin{align*}
\left|\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_n(v) -\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)}c_{n,H}(v)\right| \leq \frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} \left|c_n(v)-c_{n,H}(v)\right| =\frac{X_2}{N_n(k)}.
\end{align*}
In the third difference, we can simplify:
\begin{align*}
\left|\frac{1}{N_n(k)}\sum_{v \in V_{n,H}(k)} c_{n, H}(v)-c(k;\GPoH)\right| &= \left|\sum_{v\in V_{n,H}(k)} c_{n,H} \left(\frac{1}{N_n(k)}-\frac{1}{N_{n,H}(k)}\right)\right| \\
&= \frac{|N_{n,H}(k)-N_n(k)|}{N_n(k)N_{n,H}(k)}\sum_{v \in V_{n,H}(k)}c_{n,H}(k).
\end{align*}
Thus, we end up with the upper bound, using that $c_{n,H}(k) \leq 1$ and $N_{n,H}(k) = |V_{n,H}(k)|$,
\begin{align*}
|c(k;\GPo)-c(k;\GPoH)| \leq \frac{X_1}{N_n(k)}+\frac{X_2}{N_n(k)}+\frac{X_3}{N_n(k)}\frac{\sum_{v \in V_{n,H}(k)}c_{n,\leq H}(k)}{N_{n,H}(k)} \leq \frac{X_1+X_2+X_3}{N_n(k)}.
\end{align*}
Now using that a.a.s. $N_n(k) \geq a_1(k)n$ gives that a.a.s.
\begin{align*}
|c(k;\GPo)-c(k;\GPoH)| \leq \frac{X_1+X_2+X_3}{a_1(k)n}.
\end{align*}

By Lemma~\ref{lem:n1}, we have that $\E X_1 = \E X_3 = o(n)$ and by Lemma~\ref{lem:n2}, we have that $\E X_2 = o(n)$. We conclude that $\E |c(k;\GPo)-c(k;\GPoH)| = o(1)$. From this, convergence in probability follows by Markov's inequality: for $\epsilon >0$, $\Pee(|c(k;\GPo)-c(k;\GPoH)| \geq \epsilon) \leq \frac{\E|c(k;\GPo)-c(k;\GPoH)|}{\epsilon} = o(1)$.
\end{proof}
\subsubsection{Coupling}
\begin{lemma}\label{lem:coup}
The clustering function in the truncated Poissonized KPKVB model $\GPoH$ and in the truncated box model $\GboxH$ agree, i.e. $c(k;\GPoH) = c(k;\GboxH)$.
\end{lemma}
\begin{proof}
As all vertices involved have height $\leq H$ resp. radial coordinate $\geq R-H$ and as $H\leq \frac{R}{4}$, all vertices involved (in the disk)  will have radial coordinate $\geq \frac{3}{4}R$, therefore by Lemma~\ref{lem:coupling_edges}, edges exist between such vertices in the disk if and only if they exist in the box. So the two graphs are isomorphic and hence the clustering functions agree.
\end{proof}

@@

By the Campbell-Mecke formula, we have
\begin{align*}
\E X = \frac{1}{n}\int_{\Rcal_H} \E h(V,v) \mu(dv) = \int_0^H \E h(V,(0,y))\alpha e^{-\alpha y} dy.
\end{align*}
Then, we condition on the additional vertex $(0,y) \in \Rcal_H$ having degree $k$,
\begin{align*}
\E X = \int_0^H \E[h(V,(0,y))|D_{G_{box,H}}(0,y)=k]\Pee(D_{G_{box,H}}(0,y)=k)\alpha e^{-\alpha y}dy.
\end{align*}
We observe that for $H \rightarrow \infty$, the integrand converges point-wise for $y \in (0,\infty)$ to $P(y)\rho(y,k)\alpha e^{-\alpha y}$ and is always bounded by the integrable function $\alpha e^{-\alpha y}$.
Therefore, by the theorem of dominated convergence, as $H \rightarrow \infty$,
\begin{align*}
\E X \rightarrow \int_0^\infty P(y)\rho(y,k)\alpha e^{-\alpha y}dy.
\end{align*}
Note that for $X^2$, the summation can be split into off-diagonal (with $v_1 \not = v_2$) and diagonal (with $v_1=v_2$) terms, i.e. write $X^2 = \frac{1}{n^2}(Y_1+Y_2)$ where 
$$Y_1 = \sum_{v_1 \not = v_2 \in V(G_{box,H})} h(V(G_{box,H}),v_1)h(V(G_{box,H}),v_2)$$ and $Y_2 = \sum_{v_1 \in V(G_{box,H})} h(V(G_{box,H}),v_1)^2$. As $h \leq 1$, $Y_2 \leq n X\leq n$ and so, $\E Y_2 \leq n$, we have by the Campbell-Mecke formula,
\begin{align*}
\E Y_1 = \int_{\Rcal_H} \int_{\Rcal_H} \E[h(V(G_{box,H}),v_1)h(V(G_{box,H}),v_2)]\mu(dv_1)\mu(dv_2).
\end{align*}
If $v_1=(x_1,y_1)$ and $v_2=(x_2,y_2)$ have horizontal distance $|x_1-x_2|>2e^H$, their neighbourhood balls are disjoint w.r.t. the connection rule of the truncated box model $\GboxH$. To see this, assume that $v_3=(x_3,y_3)$ is adjacent to both $v_1$ and $v_2$. Then $|x_1-x_3|\leq e^{\frac{1}{2}(y_1+y_3)}$ and $|x_2-x_3|\leq e^{\frac{1}{2}(y_2+y_3)}$. As $y_1, y_2, y_3 \leq H$, it follows that $|x_1-x_3| \leq e^H$ and $|x_2-x_3|\leq e^H$, which implies that $|x_1-x_2| \leq |x_1-x_3|+|x_2-x_3|\leq 2e^H$, contradiction.

Due to the disjointedness of the neighbourhood balls and the properties of the Poisson process $h(V,v_1)$ and $h(V,v_2)$ are independent. The horizontal distance is uniform in $[0,\frac{\pi n}{2\nu}]$. Therefore, the probability that two vertices have horizontal distance $\leq 2e^H$ is $\leq \frac{4\nu e^H}{\pi n} = o(1)$. Writing $d_H$ for the horizontal distance, we can continue the calculation of the second moment,
\begin{align*}
\E Y_1 &= \int_{} \int_{\substack{\Rcal_H\times \Rcal_H:\\ |x_1-x_2|>2 e^H}} \E[ h(V,v_1)]\E[h(V,v_2)]\mu(dv_1)\mu(dv_2)\\
&\qquad+\int_{} \int_{\substack{\Rcal_H\times \Rcal_H:\\ |x_1-x_2|\leq2 e^H}} \E[ h(V,v_1)]\E[h(V,v_2)]\mu(dv_1)\mu(dv_2) \\
&= (1+o(1))\int_{} \int_{\Rcal_H\times \Rcal_H} \E[ h(V,v_1)]\E[h(V,v_2)]\mu(dv_1)\mu(dv_2) \\
&= (1+o(1)) (n\E X)^2.
\end{align*}
It follows that $\E X^2 = (1+o(1))(\E X)^2$, which implies $Var(X) \rightarrow 0$ using that $\E X$ tends to a constant (see above). We conclude
$$X \xrightarrow[n\rightarrow\infty]{\Pee} \int_0^\infty P(y)\rho(y,k)\alpha e^{-\alpha y}dy.$$
In a similar way, by the Campbell-Mecke formula,
\begin{align*}
\E Z =\frac{1}{n} \int_{\Rcal_H} \Pee(D_{G_{box,H}}(v)=k) \mu(dv) = \int_0^H \Pee(D_{G_{box,H}}(0,y)=k) \alpha e^{-\alpha y}dy.
\end{align*}
By the theorem of dominated convergence, it follows that
\begin{align*}
\E Z \rightarrow \int_0^\infty \rho(y,k)\alpha e^{-\alpha y}dy.
\end{align*}
The statement that $Var(Z) \rightarrow 0$ follows by the same steps as for $X$, just using $h(V,v) = \1_{\{D_{\GboxH}(v)=k\}}$ instead. Hence,
\begin{align*}
\E Z \xrightarrow[n\rightarrow\infty]{\Pee} \int_0^\infty \rho(y,k)\alpha e^{-\alpha y}dy.
\end{align*}
Finally note that 
\begin{align*}
\gamma(k) = \frac{\int_0^\infty P(y)\rho(y,k)\alpha e^{-\alpha y}dy}{\int_0^\infty \rho(y,k)\alpha e^{-\alpha y}dy}.
\end{align*}
